---
title: "Iris Feasibility Report"
author: "Christina H Thai"
date: "5/25/2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load Dataset
Installing necessary librarys
```{r}
library(e1071)
library(caTools)
library(ggplot2)
library(gridExtra)
library(grid)
```

Attaching the data and extracting the predictors into x and species classification to y
```{r}
dat= iris
```
#Splitting Data into Training and Test Set
We seek to split the data into two sets: the training set, which we will build our model with, and the test set, which we will test our model with. We want the training set to be 80% of our data (of each species). This leaves our training set to have 120 subjects and 30 subjects in our test set.
```{r}
split = sample.split(dat$Species, SplitRatio = .8)
training_set = subset(dat, split == TRUE)
test_set = subset(dat, split == FALSE)

nrow(training_set)
```

#Exploratory Visualization
Using ggplot and GGally, we can look ,ore closely at the parameters and judge if a good model can be created or not. 
```{r}
# Let's plot all the variables in a single visualization that will contain all the boxplots


BpSl <- ggplot(iris, aes(Species, Sepal.Length, fill=Species)) + 
        geom_boxplot()+
        scale_y_continuous("Sepal Length (cm)", breaks= seq(0,30, by=.5))+
        theme(legend.position="none")



BpSw <-  ggplot(iris, aes(Species, Sepal.Width, fill=Species)) + 
          geom_boxplot()+
          scale_y_continuous("Sepal Width (cm)", breaks= seq(0,30, by=.5))+
          theme(legend.position="none")



BpPl <- ggplot(iris, aes(Species, Petal.Length, fill=Species)) + 
        geom_boxplot()+
        scale_y_continuous("Petal Length (cm)", breaks= seq(0,30, by=.5))+
        theme(legend.position="none")
        


BpPw <-  ggplot(iris, aes(Species, Petal.Width, fill=Species)) + 
        geom_boxplot()+
        scale_y_continuous("Petal Width (cm)", breaks= seq(0,30, by=.5))+
        labs(title = "Iris Box Plot", x = "Species")



# Plot all visualizations
grid.arrange(BpSl  + ggtitle(""),
             BpSw  + ggtitle(""),
             BpPl + ggtitle(""),
             BpPw + ggtitle(""),
             nrow = 2,
             top = textGrob("Sepal and Petal Box Plot", 
                            gp=gpar(fontsize=15))
)
```






```{r}
pdf("data_visualize_hist.pdf")
# Sepal length 
HisSl <- ggplot(data=iris, aes(x=Sepal.Length))+
  geom_histogram(binwidth=0.2, color="black", aes(fill=Species)) + 
  xlab("Sepal Length (cm)") +  
  ylab("Frequency") + 
  theme(legend.position="none")+
  ggtitle("Histogram of Sepal Length")+
  geom_vline(data=iris, aes(xintercept = mean(Sepal.Length)),linetype="dashed",color="grey")


# Sepal width
HistSw <- ggplot(data=iris, aes(x=Sepal.Width)) +
  geom_histogram(binwidth=0.2, color="black", aes(fill=Species)) + 
  xlab("Sepal Width (cm)") +  
  ylab("Frequency") + 
  theme(legend.position="none")+
  ggtitle("Histogram of Sepal Width")+
  geom_vline(data=iris, aes(xintercept = mean(Sepal.Width)),linetype="dashed",color="grey")


# Petal length
HistPl <- ggplot(data=iris, aes(x=Petal.Length))+
  geom_histogram(binwidth=0.2, color="black", aes(fill=Species)) + 
  xlab("Petal Length (cm)") +  
  ylab("Frequency") + 
  theme(legend.position="none")+
  ggtitle("Histogram of Petal Length")+
  geom_vline(data=iris, aes(xintercept = mean(Petal.Length)),
             linetype="dashed",color="grey")

# Petal width
HistPw <- ggplot(data=iris, aes(x=Petal.Width))+
  geom_histogram(binwidth=0.2, color="black", aes(fill=Species)) + 
  xlab("Petal Width (cm)") +  
  ylab("Frequency") + 
  theme(legend.position="right" )+
  ggtitle("Histogram of Petal Width")+
  geom_vline(data=iris, aes(xintercept = mean(Petal.Width)),linetype="dashed",color="grey")


# Plot all visualizations
grid.arrange(HisSl + ggtitle(""),
             HistSw + ggtitle(""),
             HistPl + ggtitle(""),
             HistPw  + ggtitle(""),
             nrow = 2,
             top = textGrob("Iris Frequency Histogram", 
                            gp=gpar(fontsize=15))
)


dev.off()

```

We can clearly see from the Histograms of Petal.length and Petal.width that we can clearly seperate out Setosa species with very high confidence.

However, Versicolor and Virginica Species are overlapped. If we look at the scatterplot of Sepal.Length vs Petal.Length and Petal.Width vs Petal.Length, we can distintly see a seperator that can be draw between the groups of Species.

Looks like we can just use Petal.Width and Petal.Length as parameters and come with a good model. SVM seems to be a very good model for this type of data. Lets create two model, one contains all parameter and second contain just Petal.Width and Petal.Length as parameter and compare their individual performances.

#Linear Discriminant Analysis

##Model Fit and Summary
```{r}
library(MASS)
lda_model = lda(Species ~ Petal.Length + Petal.Width, dat)
```

##Model Prediction
```{r}
pred1 <- predict(lda_model,type = "response", newdata = test_set[-5])
system.time(pred1 <- predict(lda_model,type = "response", newdata = test_set[-5]))
```

##Accuracy
```{r}
table(test_set[,5],lda_prediction = pred1$class)
```



#Neural Networks
##Model Fit and Summary
```{r, warning = FALSE}
library(neuralnet)
NN_model <- neuralnet(Species ~ Petal.Length + Petal.Width, data = training_set, hidden=c(10,10), rep = 5, err.fct = "ce", linear.output = F, lifesign = "minimal", stepmax = 1000000, threshold = 0.001)

plot(NN_model, rep = "best")
```


##Model Prediction
```{r}
pred2 <- compute(NN_model, test_set[-5])
                 
system.time(pred2 <- compute(NN_model, test_set[-5]))

```

##Accuracy
```{r}
idx <- apply(pred2$net.result, 1, which.max)
predicted2 <- c('setosa', 'versicolor', 'virginica')[idx]
table(test_set[,5],predicted2)

```


#SVM Classification

#Model Fit and Summary
Create the SVM model and show summary
```{r}
svm_model <- svm(Species ~ Petal.Length + Petal.Width, data=training_set)
summary(svm_model)
```

##Model Prediction
Prediction and executtion time in R.
```{r}
pred3 <- predict(svm_model,type = "response", newdata = test_set[-5])
system.time(pred3 <- predict(svm_model,type = "response", newdata = test_set[-5]))
```

##Accuracy
Confusion matrix result from prediction, 
```{r}
table(test_set[,5],SVM_Prediction = pred3)
mis <- 2/30
```
Misclassification error is `r mis`


